---
title: "Visual Analysis"
author: "Shardendu Jha"
date: "11/14/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Part-1
# Excercise - 1

```{r }

#Loading the libaries

library(ggplot2)
library(ggthemes)
library(nlme)
library(gganimate)
library(gapminder)
library(ggExtra)
library(psych)
library(reshape2)
library(dplyr)
library(nycflights13)
library(ggcorrplot)
library(waffle)
library(tidyr)
library(scales)
library(ggalt)
library(data.table)
library(extrafont)
library(lubridate)
library(DT)
library(grid)
library(gridExtra)
library(prettydoc)
library(devtools)
library(tidyverse)
library(ggdark)
library(here)
library(png)
library(gifski)
library(forcats)
library(tufte)
library(colorspace)
library(viridisLite)
library(Zelig)
library(formatR)
library(DiagrammeR)
library(xaringan)
library(ggridges)
library(GGally)

```
```{r }

#Defining the general colors to avoid hard coding  
fill_color = 'darkred'
decoration_color = 'cyan4'
main_color = 'darkolivegreen'
sub_color = 'orange'

# customize theme
my_first_theme <- theme_gray() + theme(
  
  plot.title=element_text(size=20,face="bold", family="American Typewriter",color=decoration_color,
                          hjust=0.5,
                          lineheight=1.2),  
  plot.subtitle=element_text(size= 15, family="American Typewriter",color=decoration_color,
                             face="bold",
                             hjust=0.5), 
  plot.caption=element_text(size=15),  
  plot.background=element_blank(), 
  legend.key=element_blank(),
  legend.title=element_blank(),
  panel.grid.major=element_blank(), 
  panel.grid.minor=element_blank(),
  panel.border=element_blank(),
  legend.position=c(0.815, 0.27),
  panel.background=element_blank(),
  strip.background=element_blank(),
  strip.text =element_text(size=12, color=decoration_color),
  axis.title.x=element_text(vjust=10,size=15), 
  axis.ticks =element_blank(),
  axis.line =element_line(colour=decoration_color,size=0.3, linetype="dashed"),
  axis.text = element_text(size = 10,angle = 30,vjust = .5),  
  
)
```
# Excercise - 2
```{r }

# simple chart
ggplot(gapminder, aes(gdpPercap)) + 
  geom_freqpoly(color = sub_color)

#display with bin deatils
ggplot(gapminder, aes(gdpPercap)) + 
  geom_freqpoly(colour = sub_color, binwidth = 0.8)
#Changing the bin width (less details)
ggplot(gapminder, aes(gdpPercap)) + 
  geom_freqpoly(colour = sub_color, binwidth = 10) 

# display visual encoding as a color
ggplot(gapminder, aes(gdpPercap, colour = continent)) +
  geom_freqpoly() +
  scale_color_manual(values=c('red','orange','green','blue','purple'))

# display with range value
ggplot(gapminder, aes(gdpPercap, colour = continent)) +
  geom_freqpoly() +
  scale_color_manual(values=c('red','orange','green','blue','purple'))
  xlim(20000, 40000)

```
# Excercise - 3
```{r }
# check with histogram chart
ggplot(gapminder, aes(gdpPercap)) + 
  geom_histogram(fill= sub_color, binwidth =10)

# check range with xrange with zoom
ggplot(gapminder, aes(gdpPercap)) + 
  geom_histogram(fill=sub_color, binwidth =10) + 
  xlim(30000, 60000)

#Histogram for different cut options 
ggplot(gapminder, aes(gdpPercap, fill = continent)) +
  geom_histogram(position = "dodge") + 
  scale_fill_manual(values=c('red','orange','green','blue','purple'))

#Changing the bin options
ggplot(gapminder, aes(gdpPercap, fill = continent)) +
  geom_histogram(position = "dodge",binwidth = 10) + 
  scale_fill_manual(values=c('red','orange','green','blue','purple'))

# two diffenrent chart comes together 

ggplot(gapminder, aes(gdpPercap, fill = continent)) +
  geom_histogram(colour=decoration_color, fill = decoration_color, alpha = 0.2, size =0) +
  geom_freqpoly()+ 
  scale_colour_manual(values=c('red','orange','green','blue','purple'))

# two different chart comes together with range
ggplot(gapminder, aes(gdpPercap, colour = continent)) +
  geom_histogram(colour=decoration_color, fill = decoration_color, alpha = 0.2, size =0) +
  geom_freqpoly()+ 
  scale_colour_manual(values=c('red','orange','green','blue','purple'))
  xlim(25000,75000)

ggplot(gapminder, aes(gdpPercap)) + 
  geom_histogram(binwidth = 2,fill='blue', colour='black') + 
  facet_wrap(. ~ continent)

# change color of the chart
ggplot(gapminder, aes(gdpPercap, fill = continent)) + 
  geom_histogram(binwidth = 2,colour = 'green') +
  scale_colour_manual(values=c('red','orange','green','blue','purple'))
  facet_wrap(. ~ continent)


```
# Excercise - 4
```{r }

# Simple boxplot
ggplot(gapminder, aes(lifeExp,gdpPercap)) +
  geom_boxplot(colour=main_color)

# Boxplot using to numeric variable
ggplot(gapminder, aes(lifeExp,gdpPercap)) +
  geom_boxplot(aes(group = cut_width(gdpPercap, 10)), color='red')

# Smaller groups 
ggplot(gapminder, aes(lifeExp,gdpPercap)) +
  geom_boxplot(aes(group = cut_width(gdpPercap, 5)), color='red')

# Without outliers 
ggplot(gapminder, aes(lifeExp,gdpPercap)) +
  geom_boxplot(aes(group = cut_width(gdpPercap, 5)), color='red',outlier.alpha=0)

# encoding for the outliers
ggplot(gapminder, aes(lifeExp,gdpPercap)) +
  geom_boxplot(aes(group = cut_width(gdpPercap, 5)), color='red',
               color='blue',
               outlier.alpha = 0.5, 
               outlier.shape = 22, 
               outlier.color=decoration_color)

```
# Excercise - 5
```{r }
#Simple density chart
ggplot(gapminder, aes(gdpPercap)) +  
  geom_density(fill = main_color, color = NA)

#Multiple density chart 
ggplot(gapminder, aes(gdpPercap, group = continent, fill = continent)) +
  geom_density(adjust = 1.5 , color = NA) + 
  scale_colour_manual(values=c('red','orange','green','blue','purple'))  

#Multiple density chart, with using one color and transparency 
ggplot(gapminder, aes(gdpPercap, group = continent, fill = continent)) +
  geom_density(adjust = 1.5 , color = NA, fill='purple', alpha =0.1) 

#Small multiple density for carat by the different cuts 
ggplot(gapminder, aes(gdpPercap, group = continent, fill = continent)) + 
  geom_density(color = NA) +
  scale_colour_manual(values=c('red','orange','green','blue','purple')) +   
  facet_wrap(. ~ continent) 


```
# Excercise - 6
```{r }
#Basic scatter plot 
ggplot(gapminder, aes(x=pop, y=gdpPercap)) + 
  geom_point(color='red')

#Basic scatter plot - adjusting the size 
ggplot(gapminder, aes(x=pop, y=gdpPercap)) +  
  geom_point(size=0.5, color='red') 

#Basic scatter plot - adjusting the opacity
ggplot(gapminder, aes(x=pop, y=gdpPercap)) +  
  geom_point(alpha=0.3, color='red')

#Basic scatter plot changing the Y limits
ggplot(gapminder, aes(x=pop, y=gdpPercap)) +  
  geom_point(alpha=0.3, color='red') 
  ylim(836., 978.)

#Axis labeling depending on the quantiles 
ggplot(gapminder, aes(x=pop, y=gdpPercap)) + 
  geom_point(size=0.7, alpha=0.7, color='red')  + 
  scale_x_continuous(breaks = round(as.vector(quantile(gapminder$pop)), digits = 1))+
  scale_y_continuous(breaks = round(as.vector(quantile(gapminder$gdpPercap)), digits = 1)) 

#Adding price as another visual encoding using a colour code 
ggplot(gapminder, aes(x=pop, y=gdpPercap, colour = pop)) +
  geom_point(size=0.7, alpha=1)+
  scale_colour_gradient(low = sub_color, high = 'red') 


#Adding a trend line 
ggplot(gapminder, aes(x=pop, y=gdpPercap)) +
  geom_point(color=main_color, size=0.8, alpha=0.2)+
  stat_smooth(color=decoration_color)

#Small multiples - one variable
ggplot(gapminder, aes(x=pop, y=gdpPercap)) +
  geom_point(color=main_color, size=0.8, alpha=0.2) +
  facet_wrap( ~ continent, ncol=2) +
  stat_smooth(color=decoration_color) 

#Small multiples - one variable with free scale 
ggplot(gapminder, aes(x=pop, y=gdpPercap)) +
  geom_point(color=main_color, size=0.8, alpha=0.2) +
  facet_wrap( ~ continent, ncol=2, scales = "free") +
  stat_smooth(color=decoration_color)


```
# Excercise - 7
```{r }

#Density
pp <- ggplot(gapminder, aes(x=pop, y=gdpPercap)) +
  geom_point(color=main_color) + 
  theme(axis.title=element_blank(), axis.text=element_blank())

ggMarginal(pp, type = "density", fill = main_color, alpha=1, color='transparent')

#Box-plot  

ggMarginal(pp, type = "boxplot", size=30, fill=main_color)

#Histogram  

ggMarginal(pp, type = "histogram", size=20, fill=main_color)


```
# Excercise - 8

#Simple beewswarm
ggplot(gapminder, aes(x=gdpPercap, y=continent)) + 
  geom_quasirandom(size=3, alpha=0.7, color=main_color, groupOnX=FALSE)

#Adding color
ggplot(gapminder, aes(x=gdpPercap, y=continent, colour=continent)) + 
  geom_quasirandom(size=3, alpha=0.7, groupOnX=FALSE) +
  scale_color_manual(values=c('red','orange','green','blue','purple'))

#Adding color
ggplot(gapminder, aes(x=gdpPercap, y=continent, colour=continent)) + 
  geom_quasirandom(alpha=0.7, groupOnX=FALSE, method = "smiley") +
  scale_color_manual(values=c('red','orange','green','blue','purple'))

```{r }

#Simple beewswarm
ggplot(gapminder, aes(x=gdpPercap, y=continent)) + 
  geom_quasirandom(size=3, alpha=0.7, color=main_color, groupOnX=FALSE)

#Adding color
ggplot(gapminder, aes(x=gdpPercap, y=continent, colour=continent)) + 
  geom_quasirandom(size=3, alpha=0.7, groupOnX=FALSE) +
  scale_color_manual(values=c('red','orange','green','blue','purple'))

#Adding color
ggplot(gapminder, aes(x=gdpPercap, y=continent, colour=continent)) + 
  geom_quasirandom(alpha=0.7, groupOnX=FALSE, method = "smiley") +
  scale_color_manual(values=c('red','orange','green','blue','purple'))


```
# Excercise - 9
```{r }
#Aggregation through hexagonal binning - defining the number of bins 
ggplot(gapminder, aes(x=pop, y=gdpPercap)) +
  geom_hex(bins=60, alpha =0.8)+
  xlim(50000, 90000)
  scale_fill_gradient(low=main_color, high=sub_color) 

#Aggregation through hexagonal binning - logarithmic scaling 
ggplot(gapminder, aes(x=pop, y=gdpPercap)) +
  geom_hex(alpha = 0.6) +
  scale_x_log10(breaks = round(as.vector(quantile(gapminder$pop)), digits = 1))+
  scale_y_log10(breaks = round(as.vector(quantile(gapminder$gdpPercap)), digits = 1))+
  scale_fill_gradient(low=main_color, high=sub_color) 


```
# Excercise - 10
```{r }
#General scatter plot  
ggplot(gapminder, aes(pop, lifeExp)) +
  geom_point(colour = 'orange')   

#Apply a log scale to the X axis postiion 
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(colour = 'blue') +
  scale_x_log10()  

#Apply a a linear transformation to the Y axis position with limits 
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(colour = main_color) +
  scale_x_log10() +
  scale_y_continuous(limits = c(0, 50))
#Apply a a linear transformation to the Y axis position with defining the breaks 
ggplot(gapminder, aes(pop, lifeExp)) +
  geom_point(colour = main_color) +
  scale_x_log10() +
  scale_y_continuous(breaks = c(0, 20, 25,40, 45,55, 60, 80))

#Add labels  
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(colour = main_color) +
  scale_x_log10() +
  scale_y_continuous(breaks = c(30, 40, 60), label = c("low", "center", "high"))

#Change the Y scale in reverse 
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(colour = main_color) +
  scale_x_log10() +
  scale_y_reverse()

#Add another visual encoding size 
ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) +
  geom_point(colour = main_color) +
  scale_x_log10()  +
  scale_size()  

#Apply a scale rage to the variable size  
ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, alpha=pop)) +
  geom_point(colour = sub_color) +
  scale_x_log10()  +
  scale_size(range = c(2, 12))  

#Add another visual encoding color 
ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop,  colour = continent)) +
  geom_point() +
  scale_x_log10()  +
  scale_size(range = c(2, 12)) +
  scale_color_manual(values=c('red','orange','green','blue','purple'))

#Apply another scale to color  
ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop,  color = continent)) +
  geom_point() +
  scale_x_log10()  +
  scale_size(range = c(2, 12)) +
  scale_colour_manual(values = continent_colors) +
  scale_color_manual(values=c('red','orange','green','blue','purple'))

#Change to facet   
ggplot(gapminder, aes(gdpPercap, lifeExp,  colour = continent)) +
  geom_point() +
  scale_x_log10()  +
  scale_size(range = c(2, 12)) +
  scale_colour_manual(values = continent_colors) +
  facet_wrap(~continent) +
  scale_color_manual(values=c('red','orange','green','blue','purple'))

#Adding transparency    
ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop,  colour = continent)) +
  geom_point(alpha=0.2) +
  scale_x_log10()  +
  scale_size(range = c(2, 12)) +
  scale_colour_manual(values = continent_colors) +
  facet_wrap(~continent) +
  scale_color_manual(values=c('red','orange','green','blue','purple'))



```
# Excercise - 11
```{r }

#Check the data 
names(economics)
head(economics, n=10)
str(economics)
summary(economics)

#1 General plot definition and theme saved in a variable
e <-ggplot(economics, aes(date, pop))

#Showing the different geoms for time data 
e + geom_point() + ggtitle("point")
e + geom_bar(stat = "identity") + ggtitle("bar")
e + geom_area() + ggtitle("area")
e + geom_line() + ggtitle("line") 

#2 Showing trends for different categories 

#Check the data 
names(economics_long)
head(economics_long, n=10)
str(economics_long)
summary(economics_long)

#Check the options 
?geom_line()

#All in one chart, what is the problem?
ggplot(economics_long, aes(date, value, colour = variable)) +
  geom_line() 

#Changing to a small multiple 
ggplot(economics_long, aes(date, value)) +
  geom_line(colour=main_color) + 
  facet_wrap(~variable, ncol=5) 

#Changing to a small multiple - adding a reference line to make them comparable
ggplot(economics_long, aes(date, value)) +
  geom_line(colour=main_color) + 
  facet_wrap(~variable, ncol=2) +
  geom_hline(yintercept = 0.5, linetype = 2) 

#Changing to a small multiple - adding a reference line to make them comparable
ggplot(economics_long, aes(date, value)) +
  geom_line(colour=main_color) + 
  facet_wrap(~variable, ncol=4) +
  geom_smooth(se=FALSE, colour= sub_color, size=0.5)


#Group element in line and other charts 

#Problem with line charts 

#Check the data 
names(Oxboys)
head(Oxboys, n=10)
str(Oxboys)
summary(Oxboys)

# A single line tries to connect all the observations 
h<- ggplot(Oxboys, aes(age, height)) 
h + geom_line() 

#Grouping the observation by the variables subject
h1 <- ggplot(Oxboys, aes(age, height, group=Subject))
h1 + geom_line()

#Introducing the smooth element 

# Using the group aesthetic with both geom_line() and geom_smooth()
# groups the data the same way for both layers
h1 + geom_line() + 
  geom_smooth(aes(), colour = 'purple', size = 0.5, method = "lm", se = FALSE)

# Changing the group aesthetic for the smoother layer
# fits a single line of best fit across all boys
h1 + geom_line() +
  geom_smooth(aes(group = 1), colour = 'purple', size = 3, method = "lm", se = TRUE)

# Adding a confidence intervall
h1 + geom_line() +
  geom_smooth(aes(group = 1), colour = 'purple', size = 0.5, method = "lm", se = TRUE)

# Now we combine a box-plot with the line chart 
h2 <- ggplot(Oxboys, aes(Occasion, height))
h2 + geom_boxplot() + geom_line()

# We can add the a line chart again for all the subjects  
h2 + geom_boxplot() + geom_line(aes(group = Subject), size=0.3, colour='purple')

# We can add the a line chart grouped 
h2 + geom_boxplot() + geom_smooth(aes(group = 1), method = "lm", se = FALSE, colour='purple')

# Creating a visual analytical story 
names(gapminder)
head(gapminder, n=10)
str(gapminder)
summary(gapminder)

#General trend in life expectancy
ggplot(gapminder) +
  geom_line(aes (year, lifeExp, group = country), lwd = 0.3, show.legend = FALSE, colour = main_color) +
  labs(title = "Life expectancy has increased worldwide") 

#Checking on continents 
ggplot(gapminder) +
  geom_line(aes (year, lifeExp, group = country, color= continent), lwd = 0.3, show.legend = TRUE) +
  scale_color_manual(values=c('red','orange','green','blue','purple')) + 
  labs(title = "Life expectancy has increased worldwide") 

#Introducing a small multiple to better distinguish between continents 
ggplot() +
  geom_line(data=gapminder, aes (year, lifeExp, group = country), lwd = 0.3, show.legend = FALSE, color= main_color) + 
  facet_wrap(~ continent, ncol=5, strip.position = "bottom") +
  scale_color_manual(values=c('red','orange','green','blue','purple')) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Life expectancy by continent")

# Zooming in to see only Europe 
ggplot(subset(gapminder, continent ==  "Europe")) +
  geom_line(aes(year, lifeExp, group = country), color= main_color, show.legend = FALSE) +
  labs(title = "Life expectancy in Europe - detecting an outlier") 

# Select only Europe in order to understand which country is the outlier
europe <- dplyr::filter(gapminder, continent == "Europe")

ggplot(europe, aes(year, lifeExp)) +
  geom_line(color=main_color) +
  facet_wrap(~country) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Changes in Life Expectancy by country in europe") 

#We can also show the trend as dots 
ggplot(europe, aes(year, lifeExp)) +
  geom_point(color="grey", size=2) +
  facet_wrap(~country) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Changes in Life Expectancy by country in europe") 

#Coming back to the general checking on patterns globally 

#What will be the output of this code?

#Adding a trend line - defining the method as loss 
ggplot() +
  geom_line(data=gapminder, aes (year, lifeExp, group = country), lwd = 0.3, show.legend = FALSE, color= decoration_color) + 
  facet_wrap(~ continent, ncol=5, strip.position = "bottom") +
  geom_smooth(data=gapminder, aes(year, lifeExp, group = 1), lwd = 1, method = 'loess', span = 2, se = TRUE, color = 'purple') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Life expectancy by continent including trendline")

#We can even add all data in the background by setting the variable we do the facet with to zero 
ggplot() +
  geom_line(data = transform(gapminder, continent = NULL), aes (year, lifeExp, group = country), alpha = 0.5, lwd = 0.1, colour = "white") +
  geom_line(data=gapminder, aes (year, lifeExp, group = country), lwd = 0.3, show.legend = FALSE, color= main_color) +
  geom_smooth(data=gapminder, aes(year, lifeExp, group = 1), lwd = 1, method = 'loess', span = 0.1, se = TRUE, color = 'purple') + 
  facet_wrap(~ continent, ncol=5, strip.position = "bottom") + 
  theme(strip.background = element_blank(), strip.placement = "outside") +
  theme(axis.text.x = element_blank()) +
  labs(title = "Life expectancy by continent including trendline, showing all data in the back") 

#Showing how to add a line by aggregating the data 

#Aggregating the data 
gapminderavg<-aggregate(. ~year, data=gapminder, mean, na.rm=TRUE)
head(gapminderavg, n=10)

#Make a plot with the aggregated data 
ggplot(gapminderavg) +
  geom_line(aes (year, lifeExp), lwd = 0.3, show.legend = FALSE, color = 'purple') +
  labs(title = "Total") 

#Adding this line to the general plot by using twice the geom_line with different data sets  
ggplot() +
  geom_line(data=gapminderavg, aes (year, lifeExp), lwd = 2, show.legend = FALSE, color = 'purple') +
  geom_line(data=gapminder, aes (year, lifeExp, group = country), lwd = 0.3, show.legend = FALSE, color = "white") +
  labs(title = "Total vs. all countries")

```
# Excercise - 12
```{r }
# parallel coordinate with gapminder data

#Parallel coordinates
?ggparcoord

#Check the data 
names(gapminder)
head(gapminder, n=10)
str(gapminder)
summary(gapminder)

#Simple chart
ggparcoord(gapminder, columns = 4:6, alphaLines = 0.3)

#Simple chart, adding a color code 
ggparcoord(gapminder, columns = 4:6, groupColumn = 'continent', alphaLines = 0.3) + 
  scale_color_manual(values=c('red','orange','green','blue','purple'))

#Simple chart, adding a color code 
ggparcoord(gapminder, columns = 4:6, groupColumn = 'continent', alphaLines = 0.3, boxplot = TRUE) + 
  scale_color_manual(values=c('red','orange','green','blue','purple'))

#Showing points, changing transparency and color
ggparcoord(gapminder, columns = 4:6, groupColumn = 'continent', order = "anyClass",
           showPoints = TRUE, alphaLines = 0.3)+ 
  scale_color_manual(values=c('red','orange','green','blue','purple'))

#Showing points, changing transparency and color
ggparcoord(gapminder, columns = 4:6, groupColumn = 'continent', order = "anyClass",
           showPoints = TRUE, alphaLines = 0.3)+ 
  scale_color_manual(values=c('red','orange','green','blue','purple'))

#Different way of scaling: No scaling
ggparcoord(gapminder, columns = 4:6, groupColumn = 'continent', order = "anyClass",
           showPoints = TRUE, alphaLines = 0.3, scale="globalminmax") + 
  scale_color_manual(values=c('red','orange','green','blue','purple'))

#Different way of scaling: Standardize to Min = 0 and Max = 1
ggparcoord(gapminder, columns = 4:6, groupColumn = 'continent', order = "anyClass",
           alphaLines = 0.3, scale="uniminmax") + 
  scale_color_manual(values=c('red','orange','green','blue','purple'))+   
  facet_wrap(. ~ continent) 

#Different way of scaling: Normalize univariately (substract mean & divide by sd)
ggparcoord(gapminder, columns = 4:6, groupColumn = 'continent', order = "anyClass",
           showPoints = TRUE, alphaLines = 0.3, scale="std") + 
  scale_color_manual(values=c('red','orange','green','blue','purple'))

#Different way of scaling: Standardize and center variables
ggparcoord(gapminder, columns = 4:6, groupColumn = 'continent', order = "anyClass",
           showPoints = TRUE, alphaLines = 0.3, scale="center")+ 
  scale_color_manual(values=c('red','orange','green','blue','purple'))



```
# Excercise - 13

```{r }

#Creating a subsample 
years<- filter(gapminder, year %in% c(1952, 2007)) %>% select(country, continent, year, lifeExp)
#Check the data 
names(years)
head(years, n=10)
str(years)
summary(years)

#Convert data to wide format
years2 <- spread(years, year, lifeExp)
names(years2) <- c("country", "continent", "y1952", "y2007")

#Create a simple dumbbell plot
ggplot(years2, aes(country, x=y1952, xend=y2007))+
  geom_dumbbell(colour=main_color, size_x=0, size_xend=0)

# sorted by 2007
years3 <- arrange(years2, desc(y2007))
years3$country <- factor(years3$country, level=rev(years3$country))

#Create a simple dumbbell plot
ggplot(years3, aes(country, x=y1952, xend=y2007))+
  geom_dumbbell(colour=main_color, size_x=0, size_xend=0)

#adding color for the continent
ggplot(years3, aes(country, x=y1952, xend=y2007, color=continent))+
  scale_color_manual(values=c('red','orange','green','blue','purple'))+
  geom_dumbbell(size_x=0, size_xend=0, dot_guide=FALSE, dot_guide_size=0.2, dot_guide_colour=decoration_color)

# to declutter small multiple
ggplot(years3, aes(country, x=y1952, xend=y2007, color=continent))+
  scale_color_manual(values=c('red','orange','green','blue','purple'))+
  geom_dumbbell(size_x=0, size_xend=0, dot_guide=FALSE, dot_guide_size=0.2, dot_guide_colour=decoration_color) +
  facet_wrap(. ~ continent, ncol=5) 


#Creating a subsample 
europe2 <- filter(gapminder, continent == "Asia" & year %in% c(1952, 2007)) %>% select(country, year, lifeExp)

#Checking
head(europe2, n=10)

#Convert data to wide format
europe3 <- spread(europe2, year, lifeExp)
names(europe3) <- c("country", "y1952", "y2007")

#Checking
head(europe3, n=10)

#Create a simple dumbbell plot
ggplot(europe3, aes(country, x = y1952, xend = y2007))+
  geom_dumbbell(color=main_color)  

#Create a simple dumbbell plot
ggplot(europe3, aes(country, x = y1952, xend = y2007)) +
  geom_vline(xintercept=mean(europe3$y2007), color= decoration_color, linetype = "dashed") +
  geom_dumbbell(color=main_color)  



```
# Excercise - 14
```{r }

#Waffle chart 

?waffle

#New simple dataset created 
professional <- c(`Poor` = 520, `Rich` = 110)

#A simple waffle 
waffle(
  professional, rows = 15, size = 0.2,
  colors = c('darkgreen', 'red'), legend_pos = "bottom"
)

#You can use the iron statment to create a small multiple of waffles 
iron(
  waffle(c(thing1 = 0, thing2 = 100), colors = c('darkgreen', 'red'), rows = 5, flip=TRUE),
  waffle(c(thing1 = 25, thing2 = 75), colors = c('darkgreen', 'red'), rows = 5, flip=TRUE)
)

#It's better to add the legend then separatly instead of showing it in every chart 
iron(
  waffle(c(thing1 = 0, thing2 = 100), colors = c('darkgreen', 'red'), rows = 5, keep = FALSE, legend='none'),
  waffle(c(thing1 = 25, thing2 = 75), colors = c('darkgreen', 'red'), rows = 5, keep = FALSE, legend_pos = "bottom")
)

#5. Adding the legend only to one 

iron(
  
  waffle(
    c(`Poor` = 200, `Rich` = 110), rows = 5, size = 0.2,
    colors = c('darkgreen', 'red'),
    keep = FALSE,
    title = "Finnish citizen wealth 2019", 
    legend='none'),
  
  waffle(
    c(`Poor` = 300, `Rich` = 100), rows = 5, size = 0.2,
    colors = c('darkgreen', 'red'),
    keep = FALSE,
    title = "Estonian citizen wealth 2019",
    legend ='none'),
  
  waffle(
    c(`Poor` = 20, `Rich` = 1000), rows = 5, size = 0.2,
    colors = c('darkgreen', 'red'),
    keep = FALSE,
    title = "Swedish citizen wealth 2019",
    legend_pos = "bottom")
  
)


```

# Part - 2

```{r }

# setting the path for the folder
setwd("/Users/shardendujha/Desktop/pop")

# loading the file to R
population <- read.csv("DP_LIVE_13112020140848034.csv")

# checking the data

names(population)
head(population, n=10)
str(population)
summary(population)


ggplot(population, aes(Value)) + 
  geom_freqpoly(colour = sub_color, binwidth = 0.8)

#Changing the bin width (less details)
ggplot(population, aes(Value)) + 
  geom_freqpoly(colour = sub_color, binwidth = 10)

# display visual encoding as a color
ggplot(population, aes(Value, colour = SUBJECT)) +
  geom_freqpoly() +
  scale_color_manual(values=c('red','green'))

#Histogram for different cut options 
ggplot(population, aes(Value, fill = SUBJECT)) +
  geom_histogram(position = "dodge") + 
  scale_fill_manual(values=c('red','green'))

# Boxplot using to numeric variable
ggplot(population, aes(SUBJECT,Value)) +
  geom_boxplot(aes(group = cut_width(Value, 10)), color='red')

```

